Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-08-27 etm-transform/manuscript/main.tex
What needs to be done next

    \begin{equation*}
        \adv_\text{OT-EUF-CMA}(A) \leq \frac{L+1}{\norm{\mathbb{F}}}
    \end{equation*}
\end{lemma}
\begin{proof}
    See \cite{boneh2020graduate} lemma 7.11
\end{proof}

@@

\begin{algorithm}[H]
    \caption{\texttt{Kyber.CCAKEM.Encap$^+$(\pk)}}\label{alg:kyber-ae-encap}
    \begin{algorithmic}[1]
        \State $m \leftsample \{0,1\}^{256}$
        \State $m'= H(m)$
            \Comment{Do not output system RNG directly}
        \State $(\bar{K}, k) = G(m' \Vert H(\pk))$
            \Comment{G is instantiated with SHA3-512}
        \State $r \leftsample \mathcal{R}$
        \State $c^\prime  \leftarrow \texttt{Kyber.CPAPKE.Enc}(\pk, m', r)$
            \State $t=\mac(k, c^\prime)$
        \State $K = \texttt{KDF}(\bar{K} \Vert t)$
            \Comment{\texttt{KDF} is instantiated with Shake256}
        \State $c \leftarrow (c^\prime, t)$
        \State \Return $(c, K)$
    \end{algorithmic}
\end{algorithm}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        Name & Security level & $\keygen$ & $\encap$ & $\decap$ \\
        \hline
        Kyber512 & 128 bits & 75945 & 91467& 121185 \\
        \hline
        Kyber512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
        \hline
        Kyber512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
        \hline
        Kyber512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
        \hline
        Kyber512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
        \hline
        Kyber768 & 192 bits & 129895 & 146405& 186445 \\
        \hline
        Kyber768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
        \hline
        Kyber768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
        \hline
        Kyber768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
        \hline
        Kyber768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
        \hline
        Kyber1024 & 256 bits & 194921 & 199185 & 246245 \\
        \hline
        Kyber1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
        \hline
        Kyber1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
        \hline
        Kyber1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
        \hline
        Kyber1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
        \hline
    \end{tabular}
    \caption{Performance measurements}\label{tbl:kyber-ae-perf}
\end{table}

@@

\section{Conclusions and future works}\label{sec:future-works}
We applied the ``encrypt-then-MAC'' transformation to Kyber and saw meaningful performance improvements over using de-randomization and re-encryption. Unfortunately the resulting KEM does not achieve the desired full IND-CCA2 security, because Kyber is known to be vulnerable to key-recovery plaintext-checking attack (KR-PCA) \cite{ravi2019generic}\cite{ueno2022curse}. We speculate that while Kyber with ``encrypt-then-MAC'' could not achieve the full IND-CCA2 security, it can still be safe for use in ephemeral key exchange, where each secret key is used to decrypt at most one ciphertext (the KR-PCA requires a few hundred decryption queries to recover the secret key).

>>>>

    \begin{equation*}
        \adv_\text{OT-EUF-CMA}(A) \leq \frac{L+1}{\norm{\mathbb{F}}}
    \end{equation*}
\end{lemma}
\begin{proof}
    See \cite{boneh2020graduate} lemma 7.11
    % TODO: need more detailed citations
\end{proof}

@@

\begin{algorithm}[H]
    \caption{\texttt{Kyber.CCAKEM.Encap$^+$(\pk)}}\label{alg:kyber-ae-encap}
    \begin{algorithmic}[1]
        \State $m \leftsample \{0,1\}^{256}$
        % \State $m'= H(m)$
        %     \Comment{Do not output system RNG directly}
        \State $(\bar{K}, r, k) = G(m \Vert H(\pk))$
            \Comment{G is instantiated with SHA3-768}

        \State $c^\prime  \leftarrow \texttt{Kyber.CPAPKE.Enc}(\pk, m, r)$
            \State $t=\mac(k, c^\prime)$
        \State $K = \texttt{KDF}(\bar{K} \Vert t)$
            \Comment{\texttt{KDF} is instantiated with Shake256}
        \State $c \leftarrow (c^\prime, t)$
        \State \Return $(c, K)$
    \end{algorithmic}
\end{algorithm}

@@

\begin{table}[H]
    \centering
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        Name & Security level & $\keygen$ & $\encap$ & $\decap$ \\
        \hline
        % TODO: change to ML-KEM-512/768/1024
        ML-KEM-512 & 128 bits & 75945 & 91467& 121185 \\
        \hline
        ML-KEM-512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
        \hline
        ML-KEM-512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
        \hline
        ML-KEM-512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
        \hline
        ML-KEM-512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
        \hline
        ML-KEM-768 & 192 bits & 129895 & 146405& 186445 \\
        \hline
        ML-KEM-768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
        \hline
        ML-KEM-768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
        \hline
        ML-KEM-768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
        \hline
        ML-KEM-768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
        \hline
        ML-KEM-1024 & 256 bits & 194921 & 199185 & 246245 \\
        \hline
        ML-KEM-1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
        \hline
        ML-KEM-1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
        \hline
        ML-KEM-1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
        \hline
        ML-KEM-1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
        \hline
        % TODO: also do a short comparison with Kyber Round 3
        % TODO: KeyGen, Encap, and Decap CPU cycles should be separate tables
        % TODO: also include MAC performances on different ciphertext sizes
    \end{tabular}
    \caption{Performance measurements}\label{tbl:kyber-ae-perf}
\end{table}

@@

\section{Experimental Evaluation}
We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available \cite{}. We use the Poly1305 MAC from {\sf OpenSSL} with a key of size 256 bits and a tag size of 128 bits. Our experiments were run on a desktop with a 2.3 GHz Intel Core i9 laptop (Coffee Lake) with 16 GB RAM. The codes were compiled using \texttt{clange 14.0.0}. 
%with \texttt{-std=c99 -O1 -fomit-frame-pointer flag}. 
Table~\ref{tab:ExpResults} reports the cycle counts for three different instances of Keyber, namely . The cycle counts reported in Table~\ref{tab:ExpResults} are the average of the cycle counts of 10000 executions of all algorithms. 
\begin{table}[h]
\centering
\begin{tabular}{|c|c||c|c| |c|c| |c|c|} \hline
\multicolumn{4}{|c|}{Kyber512}  & \multicolumn{2}{|c|}{Kyber512+Poly1305} & \multicolumn{2}{|c|}{Kyber512+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 1632 (or 32) & {\bf $\keygen$:} &  & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 800 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 768 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\hline 
\multicolumn{4}{|c|}{Kyber768}  & \multicolumn{2}{|c|}{Kyber768+Poly1305} & \multicolumn{2}{|c|}{Kyber768+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 2400 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 1184 & {\bf $\encap$:} & & {\bf $\encap^+$:} &  & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 1088 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\hline 
\multicolumn{4}{|c|}{Kyber1024}  & \multicolumn{2}{|c|}{Kyber1024+Poly1305} & \multicolumn{2}{|c|}{Kyber1024+GMAC} \\ \hline
\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
{\bf sk:} & 3168 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
{\bf pk:} & 1568 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
{\bf ct:} & 1568 & {\bf $\decap$:} & & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
\end{tabular}
\caption{Cycle counts in AVX2 on a Coffee Lake laptop}
\label{tab:ExpResults}
\end{table}

\begin{itemize}
\item Forward secrecy: Mutual authentication protocol using KEM, computing the number of key-excahnge per second for both Kyber and ours. 
\item Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
\item Look at OpenSSL for GMAC with no AES-256. 
\item Also setup networking experiment on AWS
\end{itemize}

\section{Conclusions and future works}\label{sec:future-works}
\emph{Comparison with Fujisaki-Okamoto transformation:} We applied the ``encrypt-then-MAC'' transformation to Kyber and saw meaningful performance improvements over using de-randomization and re-encryption. Unfortunately the resulting KEM does not achieve the desired full IND-CCA2 security, because Kyber is known to be vulnerable to key-recovery plaintext-checking attack (KR-PCA) \cite{ravi2019generic}\cite{ueno2022curse}. We speculate that while Kyber with ``encrypt-then-MAC'' could not achieve the full IND-CCA2 security, it can still be safe for use in ephemeral key exchange, where each secret key is used to decrypt at most one ciphertext (the KR-PCA requires a few hundred decryption queries to recover the secret key).
